Reply to: 2022-18044 Intentional release of native species undermines ecological stability

### 2022-12-27 rmd/reply_r2.Rmd
Update reply_r2.Rmd

---
title: "Reply to: 2022-18044 Intentional release of native species undermines ecological stability"
output:
  bookdown::pdf_document2:
   latex_engine: xelatex
   toc: false
   number_sections: false
csl: https://www.zotero.org/styles/pnas
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
mainfont: "Times New Roman"
---

@@

This is a revision of "2022-18044 *Intentional release of native species undermines ecological stability.*" Thank you for allowing us to revise our manuscript. We have refined our analysis and text in light of comments from reviewer 3. We hope the manuscript is suitable for publication in PNAS.

@@

> Thank you. This is perhaps because of our unclear writing. At all sites, a combination of electrofishing and cast net was used to capture both benthic and water-column species effectively. We have clarified this in the revised MS (L 333 --). In addition, we have added some text to explain that the sampling has constant sampling effort per unit area (see **Supplementary text**). Therefore, our analysis is robust to this issue.

@@

> Thank you for this comment. We agree with this opinion; the main response variable in our analysis (CV of summed density across species) can be expressed as:
>
> $$
> CV = \frac{\sqrt{\sum \sigma^2_i + \sum \sum Cov_{ij}}}{\sum \mu_i}
> $$
>
> where $\mu_i$ and $\sigma_i$ are the mean and SD for species $i$. As such, the point raised by the reviewer is relevant for our analysis; fortunately, we have naturally accounted for it, as explained below.
>
> Our theory predicts that intentional release changes relative abundance of unenhanced species (and some species are competitively excluded as part of this process). The relative community composition & species richness are therefore the pathways through which intentional release affects community stability (see Theoretical predictions). Consequently, the effect of community composition is naturally integrated into the effect of intentional release, statistically speaking.
>
> Please note that, in our causal scenario, the inclusion of community composition in our regression model causes a problem called "over-control bias" (see [@arifPredictiveModelsAren2022] for details). The pathway through which intentional release affects community stability is, "intentional release (variable X1)" $\rightarrow$ "community composition (X2)" $\rightarrow$ "community stability (Y)." In this case, statistical theory suggests that including both X1 and X2 to explain Y improperly underestimates (or overlooks) the effect of X1 on Y [@arifPredictiveModelsAren2022]. Therefore, our model structure is appropriate from the perspective of statistical causal inference.
>
> In summary, we have already accounted for the point raised by the reviewer. We appreciate the input though!

@@

> Thank you for the important suggestion. We agree that different types of interactions may be important to understand the ecosystem-wide impacts of intentional release. We have clarified this point in the revised manuscript. In the meantime, we believe our research is unique and significant in that we have quantified "long-term" impact at the community level. As the reviewer suggested, there is some discussion about the community-level effects of intentional release (very few, though). However, none of them evaluated the long-term impact. We have improved the description of the knowledge gap addressed in this study (L51 --)

@@

> We understood that the reviewer commented on the last paragraph of "theoretical prediction." We have clarified that we are linking our results to reintroduction programs in conservation (paragraph starting at L135).

>>>>

---
title: "Reply to: 2022-18044 Intentional release of native species undermines ecological stability"
output:
  bookdown::pdf_document2:
   latex_engine: xelatex
   toc: false
   number_sections: false
csl: https://www.zotero.org/styles/ecology
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
mainfont: "Times New Roman"
---

@@

This is a revision of "2022-18044 *Intentional release of native species undermines ecological stability.*" Thank you for allowing us to revise our manuscript. We have refined our analysis and text in light of comments from reviewer 3.

We also noticed a very minor error in our codes while double-checking (improper setup in .RNG seed). We have re-ran the simulations after correction just in case.

We hope the manuscript is suitable for publication in PNAS.

@@

> Thank you. This is perhaps because of our unclear writing. At all sites, a combination of electrofishing and cast net was used to capture both benthic and water-column species effectively. We have clarified this in the revised MS (L. 333 --). In addition, we have added some text to explain that the sampling has constant sampling effort per unit area (see **Supplementary text**). Therefore, our analysis is robust to this issue.

@@

> Thank you for this comment. We agree with this opinion; the main response variable in our analysis (CV of summed density across species) can be expressed as:
>
> $$
> CV = \frac{\sqrt{\sum \sigma^2_i + \sum \sum Cov_{ij}}}{\sum \mu_i}
> $$
>
> where $\mu_i$ and $\sigma_i$ are the mean and SD for species $i$. As such, the point raised by the reviewer is relevant for our analysis; fortunately, we have naturally accounted for it, as explained below.
>
> Our theory predicts that intentional release changes relative abundance of unenhanced species (and some species are competitively excluded as part of this process). The relative community composition & species richness are therefore the pathways through which intentional release affects community stability (see Theoretical prediction). Consequently, the effect of community composition is naturally integrated into the effect of intentional release, statistically speaking.
>
> Please note that, in our causal scenario, the inclusion of community composition in our regression model causes a problem called "over-control bias" (see @arifPredictiveModelsAren2022 for details). The pathway through which intentional release affects community stability is, "intentional release (variable X1)" $\rightarrow$ "community composition (X2)" $\rightarrow$ "community stability (Y)." In this case, statistical theory suggests that including both X1 and X2 to explain Y improperly underestimates (or overlooks) the effect of X1 on Y [@arifPredictiveModelsAren2022]. Therefore, our model structure is appropriate from the perspective of statistical causal inference.
>
> In summary, we have already accounted for the point raised by the reviewer. We appreciate the input though!

@@

> Thank you for the important suggestion. We agree that different types of interactions may be important to understand the ecosystem-wide impacts of intentional release. We have clarified this point in the revised manuscript. In the meantime, we believe our research is unique and significant in that we have quantified "long-term" impact at the community level. As the reviewer suggested, there is some discussion about the community-level effects of intentional release (very few, though). However, none of them evaluated the long-term impact. We have improved the description of the knowledge gap addressed in this study (L. 51 --)

@@

> We understood that the reviewer commented on the last paragraph of "theoretical prediction." We have clarified that we are linking our results to reintroduction programs in conservation (paragraph starting at L. 135).
